\pkg{Rgbp}: An R Package for Gaussian, Poisson, and Binomial Hierarchical Modeling

### 2013-09-11 jss/article.tex
discussion done

where $\mu_{0j} = x^{T}_{j} \mathbf{\beta}$, $j=1, \ldots k$, $x_{j}$ is the $j$-th group's covariate vector ($m\times 1$), $m$ is the number of regression coefficients and both $\mathbf{\beta}$ and $A$ are unknown. Note that if there is no covariate then $x_{j}=1$ for an intercept term ($m=1$) and so $\mu_{0j}=\mu_{0}=\beta_{0}$ for all $j$, resulting in an exchangeable prior distribution. For reference, a parameter with a zero subscript, such as $\mu_{0j}$, represents a mean parameter of the prior (second-level) distribution, \emph{i.e.}, a prior mean. Based on this conjugate prior distribution it is easy to derive the corresponding posterior distribution
\begin{equation} \label{normalpost}
\mu_{j}\vert \textbf{y}, \beta, A \stackrel{ind}{\sim}\textrm{Normal}(~(1-B_{j})y_{j} + B_{j}\mu_{0j},~(1-B_{j})V_{j}~),
\end{equation}
where $B_{j}\equiv V_{j}/(V_{j} + A),~j=1, \ldots, k$, are called shrinkages. By integrating over $\mu_{j}$ for $j = 1 \ldots k$ a marginal distribution, useful for model checking (Section \ref{discussion}), can be derived
\begin{equation} \label{normalmarginal}
y_{j}\vert \beta, A \stackrel{ind}{\sim}\textrm{Normal}(\mu_{0j}, V_{j} + A). 
\end{equation}

@@

A benefit of \pkg{Rgbp} is that it produces non-random output and so results can easily be reproduced and compared across studies. In addition to being a standalone analysis tool the package can be used as an aid or a tool in a broader estimation procedure. For example, convergence of an MCMC can be increased by using the point estimates from \pkg{Rgbp} to initialize the chain.

Due to the speed and ease of use \pkg{Rgbp} can be used as a method of preliminary data analysis. The results of which may inform the statistician whether a more intensive (in terms of coding and computational time) but more precise method (such as MCMC) should be used.

@@

to initialize an MCMC to speed up time to convergence. to check the validity of the code of an MCMC program.

as a demo that a particular prior gives adequate coverage. for frequentist procedure checking.

>>>>

where $\mu_{0j} = x^{T}_{j} \mathbf{\beta}_{(m\times1)}=\beta_{0}+\beta_{1}x_{1, j} + \cdots + \beta_{m-1}x_{m-1, j}$, $j=1, \ldots k$ and $m$ is the number of regression coefficients including an intercept term. The notation $x_{j}^{T}=(1, x_{1, j}, x_{2, j}, \ldots, x_{m-1, j})$ represents the $j$-th group's $m\times 1$ covariate vector whose first element is 1 for the intercept term of regression coefficients. Both $\mathbf{\beta}_{(m\times1)}$ and the second-level variance $A$ are unknown. Note that if there is no covariate then $x_{j}=1$ just for an intercept term ($m=1$) and so $\mu_{0j}=\mu_{0}=\beta_{0}$ for all $j$, resulting in an exchangeable prior distribution. For reference, a parameter with a zero subscript, such as $\mu_{0j}$, represents a mean parameter of the prior (second-level) distribution, \emph{i.e.}, a prior mean. Based on this conjugate prior distribution it is easy to derive the corresponding posterior distribution

\begin{equation} \label{normalpost}
\mu_{j}\vert \textbf{y}, \beta, A \stackrel{ind}{\sim}\textrm{Normal}(~(1-B_{j})y_{j} + B_{j}\mu_{0j},~(1-B_{j})V_{j}~),
\end{equation}
where $B_{j}\equiv V_{j}/(V_{j} + A),~j=1, \ldots, k$, are called shrinkages. By integrating over $\mu_{j}$ for $j = 1 \ldots k$ a marginal distribution, useful for model checking (Section \ref{discussion}), can be derived
\begin{equation} \label{normalmarginal}
y_{j}\vert \beta, A \stackrel{ind}{\sim}\textrm{Normal}(\mu_{0j}, V_{j} + A). 
\end{equation}

@@

A benefit of \pkg{Rgbp} is that it produces non-random output and so results can easily be reproduced and compared across studies. In addition to being a standalone analysis tool the package can be used as an aid or a tool in a broader estimation procedure. For example, by checking the similarity of output of \pkg{Rgbp} and that of another estimation procedure (such as MCMC) the package can be used as a confirmatory tool to check whether the alternative procedure has been implemented correctly. In addition, the parameter estimates obtained via \pkg{Rgbp} can be used to initialize a MCMC thus increasing time to convergence. Note, however, that this is only valid when the same model under both procedures is assumed.

Due to the speed and ease of use \pkg{Rgbp} can be utilized as a method of preliminary data analysis. The results of which may inform the statistician whether a more intensive (in terms of coding and computational time) but more precise method (such as MCMC) should be used. As the package assumes non-informative priors the results can be used as reference point to see whether implementing a more sophisticated model, perhaps with informative priors, should be used.

@@
